---
title: "ML_Assignment"
author: "Alex Perez Nava"
date: "1/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit, it is now possible to collect a large amount of data about personal activity relatively inexpensively. The aim of this project is to predict the manner in which participants perform a barbell lift. The data comes from http://groupware.les.inf.puc-rio.br/har wherein 6 participants were asked to perform the same set of exercises correctly and incorrectly with accelerometers placed on the belt, forearm, arm, and dumbbell.

## Background
“Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).

Class A corresponds to the specified execution of the exercise, while the other 4 classes correspond to common mistakes. Participants were supervised by an experienced weight lifter to make sure the execution complied to the manner they were supposed to simulate. The exercises were performed by six male participants aged between 20-28 years, with little weight lifting experience. We made sure that all participants could easily simulate the mistakes in a safe and controlled manner by using a relatively light dumbbell (1.25kg)."


```{r include=FALSE}
library(knitr)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(corrplot)
```

## Get and process data
Get data set from the URL below. The training data set is  partitioned in two to create a Training set (70% of the data) for the modeling process and a Test set (30%) for the validations.

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from http://groupware.les.inf.puc-rio.br/har. Full source:

Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. “Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human ’13)”. Stuttgart, Germany: ACM SIGCHI, 2013.

# Set the URL for the download
```{r }
UrlTrain <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
UrlTest  <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
```

# Download the datasets
```{r }
training <- read.csv(url(UrlTrain))
testing  <- read.csv(url(UrlTest))
```

# Create a partition with the training data set
```{r }
inTrain  <- createDataPartition(training$classe, p=0.7, list=FALSE)
TrainSet <- training[inTrain, ]
TestSet  <- training[-inTrain, ]
dim(TrainSet)
```

```{r }
dim(TestSet)
```

Created data sets have 160 variables. Clean data set removing NA's. Near Zero variance (NZV) variables and ID.

# Remove variables with Nearly Zero Variance
```{r }
NZV <- nearZeroVar(TrainSet)
TrainSet <- TrainSet[, -NZV]
TestSet  <- TestSet[, -NZV]
dim(TrainSet)
```

```{r }
dim(TestSet)
```

# Remove variables that are mostly NA
```{r echo=FALSE}
AllNA    <- sapply(TrainSet, function(x) mean(is.na(x))) > 0.95
TrainSet <- TrainSet[, AllNA==FALSE]
TestSet  <- TestSet[, AllNA==FALSE]
dim(TrainSet)
```

```{r }
dim(TestSet)
```

# Remove identification only variables (columns 1 to 5)
```{r echo=FALSE}
TrainSet <- TrainSet[, -(1:5)]
TestSet  <- TestSet[, -(1:5)]
dim(TrainSet)
```

```{r }
dim(TestSet)
```

The data sets for the analysis have now 54 variables as a result of the cleaning process.

```{r }
names(TrainSet)
```

## Correlation analysis

```{r echo=FALSE}
corMatrix <- cor(TrainSet[, -54])
corrplot(corMatrix, order = "FPC", method = "circle", type = "upper", tl.cex = 0.4, col = heat.colors(100))
```

Positive correlation are shown in yellow circles and negative in red. Color intensity and size of the circle are proportional to the correlation coefficients.

## Prediction Model
We want to compare methods in order to identify the one with higher accuracy then applied it to the Test data set. The selected method will be used for the quiz predictions. The methods are: Random Forests, Decision Tree and Generalized Boosted Model. Look at the Confusion Matrix of each method.

## I.-Random Forest

model fit

```{r echo=FALSE}
set.seed(12345)
controlRF <- trainControl(method="cv", number=3, verboseIter=FALSE)
modFitRandForest <- train(classe ~ ., data=TrainSet, method="rf",trControl=controlRF)
modFitRandForest$finalModel
```

# Prediction on Test dataset
```{r echo=FALSE}
predictRandForest <- predict(modFitRandForest, newdata=TestSet)
confMatRandForest <- confusionMatrix(predictRandForest, TestSet$classe)
confMatRandForest
```

# Plot matrix results
```{r echo=FALSE}
plot(confMatRandForest$table, col = confMatRandForest$byClass, main = paste("Random Forest - Accuracy =",round(confMatRandForest$overall['Accuracy'], 4)))
```

## II.-Decision Trees

model fit

```{r echo=FALSE}
set.seed(12345)
modFitDecTree <- rpart(classe ~ ., data=TrainSet, method="class")
fancyRpartPlot(modFitDecTree)
```

# Prediction on Test data set
```{r echo=FALSE}
predictDecTree <- predict(modFitDecTree, newdata=TestSet, type="class")
confMatDecTree <- confusionMatrix(predictDecTree, TestSet$classe)
confMatDecTree
```

# Plot matrix results
```{r echo=FALSE}
plot(confMatDecTree$table, col = confMatDecTree$byClass, main = paste("Decision Tree - Accuracy =",round(confMatDecTree$overall['Accuracy'], 4)))
```

## III.-Generalized Boosted Model

model fit
```{r echo=FALSE}
set.seed(12345)
controlGBM <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
modFitGBM  <- train(classe ~ ., data=TrainSet, method = "gbm",trControl = controlGBM, verbose = FALSE)
modFitGBM$finalModel
```

# Prediction on Test data set
```{r echo=FALSE}
predictGBM <- predict(modFitGBM, newdata=TestSet)
confMatGBM <- confusionMatrix(predictGBM, TestSet$classe)
confMatGBM
```

# Plot matrix results
```{r echo=FALSE}
plot(confMatGBM$table, col = confMatGBM$byClass, main = paste("GBM - Accuracy =",round(confMatGBM$overall['Accuracy'], 4)))
```

The accuracy of the 3 regression modeling methods above are:

Random Forest : 0.9968
Decision Tree : 0.7373
GBM : 0.9861
In conclusion, the Random Forest model will be applied to predict the 20 quiz results.


predictTEST <- predict(modFitRandForest, newdata=testing)

predictTEST


